Updated results

### 2021-03-12 analysis/reports/manuscript_main.Rmd
get cts without ones that haven't run eyt

simpson_ks <- filter(sub_di, nparts > 20) %>%
  select(simpson_95_ratio_1t, Dataset, simpson_percentile, s0, n0) %>% 
  group_by(Dataset, s0, n0) %>%
  mutate(pairing = dplyr::row_number()) %>%
  ungroup() %>%
  tidyr::pivot_wider(id_cols = c(s0, n0, pairing), names_from = Dataset, values_from = c(simpson_95_ratio_1t, simpson_percentile))

@@

skewness_ks <- filter(sub_di, nparts > 20) %>%
  select(skew_95_ratio_1t, Dataset, skew_percentile, s0, n0) %>% 
  group_by(Dataset, s0, n0) %>%
  mutate(pairing = dplyr::row_number()) %>%
  ungroup() %>%
  tidyr::pivot_wider(id_cols = c(s0, n0, pairing), names_from = Dataset, values_from = c(skew_95_ratio_1t, skew_percentile))

>>>>

simpson_ks <- filter(sub_di, nparts > 40) %>%
  select(simpson_95_ratio_1t, Dataset, simpson_percentile, s0, n0) %>% 
  group_by(Dataset, s0, n0) %>%
  mutate(pairing = dplyr::row_number()) %>%
  ungroup() %>%
  tidyr::pivot_wider(id_cols = c(s0, n0, pairing), names_from = Dataset, values_from = c(simpson_95_ratio_1t, simpson_percentile))

@@

skewness_ks <- filter(sub_di, nparts > 40) %>%
  select(skew_95_ratio_1t, Dataset, skew_percentile, s0, n0) %>% 
  group_by(Dataset, s0, n0) %>%
  mutate(pairing = dplyr::row_number()) %>%
  ungroup() %>%
  tidyr::pivot_wider(id_cols = c(s0, n0, pairing), names_from = Dataset, values_from = c(skew_95_ratio_1t, skew_percentile))

### 2021-03-12 analysis/rev_prototyping/revision_results.Rmd
get cts without ones that haven't run eyt

all_di_shapes_long <- all_di  %>%
  filter(nparts > 40) %>%
  select(dat, skew_percentile, skew_percentile_excl, simpson_percentile, simpson_percentile_excl, shannon_percentile, shannon_percentile_excl, nsingletons_percentile, nsingletons_percentile_excl) %>%
  rename(skew = skew_percentile,
         skewX = skew_percentile_excl,
         simp = simpson_percentile,
         simpX = simpson_percentile_excl,
         shan = shannon_percentile,
         shanX = shannon_percentile_excl,
         nsng = nsingletons_percentile,
         nsngX = nsingletons_percentile_excl) %>%
  mutate(skewM = (skew + skewX) / 2,
         simpM = (simp+ simpX)/2,
         shanM = (shan+ shanX)/2,
         nsngM = (nsng+ nsngX)/2) %>%
  tidyr::pivot_longer(-dat, names_to = "metric", values_to = "value") %>%
  mutate(metric_name = substr(metric, 0, 4),
         metric_mod = ifelse(substr(metric, 5, 5) == "X", "excl", ifelse(substr(metric, 5,5) == "M", "mean", "incl")))

@@

all_di_shapes_high <- all_di_shapes_long %>%
  filter(metric_mod == "excl") %>%
  group_by(dat, metric_name) %>%
  summarize(proportion_high = mean(value > 95))

all_di_shapes_high %>%
  tidyr::pivot_wider(id_cols = dat, names_from = metric_name, values_from = proportion_high) %>% mutate(dir = "HIGH")

@@

```
<!-- ``` -->
<!-- ```{r} -->

<!-- ggplot(all_di, aes((real_po / sim_pos_from_best), real_po_percentile_excl, color =  real_po_percentile_excl > 95)) + -->
<!--   geom_point(alpha = .3)  -->

<!-- all_di %>% -->
<!--   mutate(rat = real_po / sim_pos_from_best, -->
<!--          dev = real_po_percentile_excl > 95) %>% -->
<!--   group_by(dev) %>% -->
<!--   summarize(minrat = min(rat), -->
<!--             maxrat = max(rat), -->
<!--             meanrat = mean(rat), -->
<!--             mindis = min(real_po - sim_pos_from_best), -->
<!--             maxdis = max(real_po - sim_pos_from_best)) -->

<!-- ggplot(all_di, aes(nparts, sim_pos_from_best)) + -->
<!--   geom_point() + -->
<!--   scale_x_log10() -->

<!-- ggplot(all_di, aes(nparts, sim_pos_from_best, color = dat)) + -->
<!--   geom_point() + -->
<!--   scale_x_log10()  + -->
<!--   scale_color_viridis_d() + -->
<!--   theme(legend.position = "top") -->

<!-- ggplot(all_di, aes(s0, n0, color =  sim_pos_from_best)) + -->
<!--   geom_point() + -->
<!--   scale_x_log10()  + -->
<!--   scale_y_log10() + -->
<!--   scale_color_viridis_c() + -->
<!--   theme(legend.position = "top") -->
<!-- ``` -->

>>>>

all_di_shapes_long <- all_di  %>%
  filter(nparts > 40) %>%
  mutate(dat = ifelse(nparts < 1000, paste0(dat, "_small"), dat)) %>%
  select(dat, skew_percentile, skew_percentile_excl, simpson_percentile, simpson_percentile_excl, shannon_percentile, shannon_percentile_excl, nsingletons_percentile, nsingletons_percentile_excl) %>%
  rename(skew = skew_percentile,
         skewX = skew_percentile_excl,
         simp = simpson_percentile,
         simpX = simpson_percentile_excl,
         shan = shannon_percentile,
         shanX = shannon_percentile_excl,
         nsng = nsingletons_percentile,
         nsngX = nsingletons_percentile_excl) %>%
  mutate(skewM = (skew + skewX) / 2,
         simpM = (simp+ simpX)/2,
         shanM = (shan+ shanX)/2,
         nsngM = (nsng+ nsngX)/2) %>%
  tidyr::pivot_longer(-dat, names_to = "metric", values_to = "value") %>%
  mutate(metric_name = substr(metric, 0, 4),
         metric_mod = ifelse(substr(metric, 5, 5) == "X", "excl", ifelse(substr(metric, 5,5) == "M", "mean", "incl")))

@@

all_di_shapes_high <- all_di_shapes_long %>%
  filter(metric_mod == "excl") %>%
  group_by(dat, metric_name) %>%
  summarize(proportion_high = mean(value > 95),
            n = dplyr::n())

all_di_shapes_high %>%
  tidyr::pivot_wider(id_cols = c(dat, n), names_from = metric_name, values_from = proportion_high) %>% mutate(dir = "HIGH")

@@

all_di_shapes_high <- all_di_shapes_long %>%
  mutate(in_fia = ifelse(dat == "fia", "FIA", "Other")) %>%
  filter(metric_mod == "excl") %>%
  group_by(in_fia, metric_name) %>%
  summarize(proportion_high = mean(value >95))

all_di_shapes_high %>%
  tidyr::pivot_wider(id_cols = in_fia, names_from = metric_name, values_from = proportion_high) %>%
  mutate(dir = "high")

all_di_shapes_low <- all_di_shapes_long %>%
  mutate(in_fia = ifelse(grepl("fia", dat), "FIA", "Other")) %>%
  filter(metric_mod == "incl") %>%
  group_by(in_fia, metric_name) %>%
  summarize(proportion_low = mean(value <5))

all_di_shapes_low %>%
  tidyr::pivot_wider(id_cols = in_fia, names_from = metric_name, values_from = proportion_low) %>%
  mutate(dir = "LOW")

## Narrowness of the expectation

```{r}
ggplot(all_di, aes(nparts, sim_pos_from_best, color = log(n0/s0))) +
  geom_point() +
  scale_x_log10() +
  scale_color_viridis_c(option = "plasma")

```{r, fig.dim = c(3,6)}

all_di <- all_di %>%
  mutate(tiny = nparts < 2000)

ggplot(all_di, aes(sim_pos_from_best)) +
geom_histogram(bins = 100) +
  facet_wrap(vars(tiny), scales = "free_y", ncol = 1)

### Shape metric

```{r}

ggplot(filter(all_di, nparts < 1e56, s0 > 2), aes(nparts, skew_95_ratio_2t, color = log(n0/s0))) +
  geom_point() +
  scale_x_log10() +
  scale_color_viridis_c(option = "plasma")

ggplot(filter(all_di, nparts < 1e56), aes(nparts, simpson_95_ratio_2t, color = log(n0/s0))) +
  geom_point() +
  scale_x_log10() +
  scale_color_viridis_c(option = "plasma")

ggplot(filter(all_di, nparts < 1e56), aes(nparts, shannon_95_ratio_2t, color = log(n0/s0))) +
  geom_point() +
  scale_x_log10() +
  scale_color_viridis_c(option = "plasma")

ggplot(filter(all_di, nparts < 1e56), aes(nparts, nsingletons_95_ratio_2t, color = log(n0/s0))) +
  geom_point() +
  scale_x_log10() +
  scale_color_viridis_c(option = "plasma")

```{r, fig.dim = c(4, 6)}

ggplot(filter(all_di), aes(sim_pos_from_best)) +
geom_histogram() +
  facet_wrap(vars(tiny), scales = "free_y", ncol = 1)
ggplot(filter(all_di, s0 > 2), aes(skew_95_ratio_2t)) +
geom_histogram() +
  facet_wrap(vars(tiny), scales = "free_y", ncol = 1)

ggplot(filter(all_di), aes(simpson_95_ratio_2t)) +
geom_histogram() +
  facet_wrap(vars(tiny), scales = "free_y", ncol = 1)

ggplot(filter(all_di), aes(shannon_95_ratio_2t)) +
geom_histogram() +
  facet_wrap(vars(tiny), scales = "free_y", ncol = 1)

ggplot(filter(all_di), aes(nsingletons_95_ratio_2t)) +
geom_histogram() +
  facet_wrap(vars(tiny), scales = "free_y", ncol = 1)
```

```{r, fig.dim = c(3, 6)}

ggplot(filter(all_di, dat  %in% c("fia", "mcdb", "misc_abund")), aes(sim_pos_from_best)) +
geom_histogram() +
  facet_wrap(vars(tiny, dat), scales = "free_y", ncol = 3)
ggplot(filter(all_di, s0 > 2, dat  %in% c("fia", "mcdb", "misc_abund")), aes(skew_95_ratio_2t)) +
geom_histogram() +
  facet_wrap(vars(tiny, dat), scales = "free_y", ncol = 3)

ggplot(filter(all_di, dat  %in% c("fia", "mcdb", "misc_abund")), aes(simpson_95_ratio_2t)) +
geom_histogram() +
  facet_wrap(vars(tiny, dat), scales = "free_y", ncol = 3)

ggplot(filter(all_di, dat  %in% c("fia", "mcdb", "misc_abund")), aes(shannon_95_ratio_2t)) +
geom_histogram() +
  facet_wrap(vars(tiny, dat), scales = "free_y", ncol = 3)

ggplot(filter(all_di, dat  %in% c("fia", "mcdb", "misc_abund")), aes(nsingletons_95_ratio_2t)) +
geom_histogram() +
  facet_wrap(vars(tiny, dat), scales = "free_y", ncol = 3)

```{r, fig.dim = c(7,7)}
library(grid)

fia_max_s = max(filter(all_di, dat == "fia")$s0)
fia_max_n = max(filter(all_di, dat == "fia")$n0)

fia = filter(all_di, dat == "fia")

small_di <- all_di %>%
  filter(s0 <= fia_max_s,
         n0 <= fia_max_n) %>%
  mutate(fia_yn = ifelse(dat == "fia", "fia", "other datasets"),
         right_corner = FALSE)

for(i in 1:nrow(small_di)) {
  if(small_di$dat[i] != "fia") {
    fia_match_s0 = filter(fia, s0 >= small_di$s0[i])

    max_n0_for_this_s0 = max(fia_match_s0$n0)

    small_di$right_corner[i] = small_di$n0[i] > max_n0_for_this_s0

  }
}

small_di <- small_di %>%
  filter(!right_corner)
not_fia <- filter(small_di, fia_yn != "fia")
fia = filter(small_di, dat == "fia")

small_di_s_n <- small_di %>%
  select(s0, n0) %>%
  distinct() %>%
  mutate(in_fia = FALSE,
         in_not_fia = FALSE)

for(i in 1:nrow(small_di_s_n)) {

  this_s0 = small_di_s_n$s0[i]
  this_n0 = small_di_s_n$n0[i]

  fia_matches = filter(fia, s0 == this_s0, n0 == this_n0)
  not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0)

  small_di_s_n$in_fia[i] = nrow(fia_matches) > 0 
  small_di_s_n$in_not_fia[i] = nrow(not_fia_matches) > 0

}

small_di_s_n <- filter(small_di_s_n, in_fia, in_not_fia)

set.seed(1977)

subsamples <- list()

for(i in 1:nrow(small_di_s_n)) {

  this_s0 = small_di_s_n$s0[i]
  this_n0 = small_di_s_n$n0[i]

  fia_matches = filter(fia, s0 == this_s0, n0 == this_n0)

  not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0)

  n_to_draw = min(nrow(fia_matches), nrow(not_fia_matches))

  fia_draw = sample.int(n = nrow(fia_matches), size = n_to_draw, replace = F)
  not_fia_draw = sample.int(n = nrow(not_fia_matches), size = n_to_draw, replace = F)

  subsamples[[i]] <- (bind_rows(fia_matches[fia_draw,], not_fia_matches[not_fia_draw, ]))

}

sub_di <- bind_rows(subsamples) %>%
  mutate(Dataset = ifelse(fia_yn == "fia", "FIA", "Other datasets"))

simpson_ks <- filter(sub_di, nparts > 40) %>%
  select(simpson_95_ratio_2t, Dataset, simpson_percentile, s0, n0) %>% 
  group_by(Dataset, s0, n0) %>%
  mutate(pairing = dplyr::row_number()) %>%
  ungroup() %>%
  tidyr::pivot_wider(id_cols = c(s0, n0, pairing), names_from = Dataset, values_from = c(simpson_95_ratio_2t, simpson_percentile))

simpson_breadth <- ks.test(simpson_ks$simpson_95_ratio_2t_FIA, simpson_ks$`simpson_95_ratio_2t_Other datasets`)

simpson_percent <- ks.test(simpson_ks$simpson_percentile_FIA, simpson_ks$`simpson_percentile_Other datasets`)
simpson_breadth
simpson_percent

skewness_ks <- filter(sub_di, nparts > 40) %>%
  select(skew_95_ratio_2t, Dataset, skew_percentile_excl, s0, n0) %>% 
  group_by(Dataset, s0, n0) %>%
  mutate(pairing = dplyr::row_number()) %>%
  ungroup() %>%
  tidyr::pivot_wider(id_cols = c(s0, n0, pairing), names_from = Dataset, values_from = c(skew_95_ratio_2t, skew_percentile_excl))

skewness_breadth <- ks.test(skewness_ks$skew_95_ratio_2t_FIA, skewness_ks$`skew_95_ratio_2t_Other datasets`)

skewness_percent <- ks.test(skewness_ks$skew_percentile_excl_FIA, skewness_ks$`skew_percentile_excl_Other datasets`)
skewness_breadth
skewness_percent

ggplot(filter(sub_di, nparts > 40), aes(real_po_percentile_excl)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn))

ggplot(filter(sub_di, nparts > 40), aes(sim_pos_from_best)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn))

ggplot(filter(sub_di, nparts > 40), aes(skew_percentile_excl)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn))

ggplot(filter(sub_di, nparts > 40), aes(skew_95_ratio_2t)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn))

ggplot(filter(sub_di, nparts > 40), aes(nsingletons_percentile_excl)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn))

ggplot(filter(sub_di, nparts > 40), aes(nsingletons_95_ratio_2t)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn))

ggplot(filter(sub_di, nparts > 20), aes(simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn))

ggplot(filter(sub_di, nparts > 40), aes(simpson_95_ratio_2t)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn))
ggplot(filter(sub_di, nparts > 40), aes(shannon_percentile)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn))

ggplot(filter(sub_di, nparts > 40), aes(shannon_95_ratio_2t)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn))

```{r}
# 
# fig_4 <- gridExtra::grid.arrange(grobs = list(
#   ggplot(filter(sub_di, nparts > 40), aes(skew_95_ratio_2t)) +
#     geom_histogram()+
#     theme_bw() +
#     facet_wrap(vars(Dataset), scales = "free_y")  +
#     xlab("Breadth index") +
#     ylab("Number of communities") +
#     ggtitle("Breadth index", subtitle = "Skewness"),
#   
#   ggplot(filter(sub_di, nparts > 40), aes(simpson_95_ratio_2t)) +
#     geom_histogram()+
#     theme_bw() +
#     facet_wrap(vars(Dataset), scales = "free_y")  +
#     xlab("Breadth index") +
#     ylab("") +
#     ggtitle("", subtitle = "Evenness"),
#   
#   ggplot(filter(sub_di, nparts > 40), aes(skew_percentile_excl)) +
#     geom_histogram()+
#     theme_bw() +
#     facet_wrap(vars(Dataset), scales = "free_y")  +
#     xlab("Percentile rank") +
#     ylab("Number of communities") +
#     ggtitle("Percentile rank", subtitle = "Skewness"),
#   
#   
#   ggplot(filter(sub_di, nparts > 40), aes(simpson_percentile)) +
#     geom_histogram()+
#     theme_bw() +
#     facet_wrap(vars(Dataset), scales = "free_y")  +
#     xlab("Percentile rank") +
#     ylab("")+
#     ggtitle("", subtitle = "Evenness")), ncol = 2, top = textGrob("Figure 4", gp = gpar(fill = "white")))
# 
# plot(fig_4)
```

```{r}

all_di <- all_di %>%
  mutate(tiny = nparts < exp(10))

ggplot(filter(all_di, nparts > 40, s0 > 2), aes(sim_pos_from_best, color = tiny)) +
  geom_density()
ggplot(filter(all_di, nparts > 40, s0 > 2), aes(skew_95_ratio_2t, color = tiny)) +
  geom_density()

ggplot(filter(all_di, nparts > 40), aes(simpson_95_ratio_2t, color = tiny)) +
  geom_density()

ggplot(filter(all_di, nparts > 40), aes(shannon_95_ratio_2t, color = tiny)) +
  geom_density()
```

```{r}

## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", "drake-cache-net.sqlite"))
cache <- storr::storr_dbi("datatable", "keystable", db)
cache$del(key = "lock", namespace = "session")
knitr::opts_chunk$set(fig.width=3, fig.height=2.5, warning = FALSE, message = FALSE)

net_summary <- readd(all_di_summary, cache = cache)

net_summary <- net_summary %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)))

ggplot(net_summary, aes(log_nparts, skew_95_ratio_2t)) +
  geom_point()
